\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

\title{Chapter 10 Exercises\\Gallian's Book on Abstract Algebra}
\author{Spencer T. Parkin}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{identity}{Identity}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{result}{Result}[section]

%\newcommand{\gcd}{\mbox{gcd}}
\newcommand{\lcm}{\mbox{lcm}}
\newcommand{\abs}{\mbox{abs}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\G}{\mathbb{G}}
\newcommand{\stab}{\mbox{stab}}
\newcommand{\aut}{\mbox{Aut}}
\newcommand{\inn}{\mbox{Inn}}
\newcommand{\orb}{\mbox{orb}}

\begin{document}
\maketitle

\section*{Lemma 1}

Let $H$ be a proper subgroup of $G$.  Then
for all $g\in G-H$ and all $h\in H$, $gh\in G-H$.

Suppose $gh=h'\in H$.  Then $g=h'h^{-1}\in H$,
which is a contradiction.  Therefore, $gh\in G-H$.

\section*{Lemma 2}

Let $N$ be a normal subgroup of a group $G$.
Then for any $g\in G$ and any $n\in N$,
there exists $n'\in N$ such that $gn=n'g$ or such that $ng=gn'$.

\section*{Lemma 3}

Let $G$ be a group and let $n$ be a positive integer.
Then the number of elements in $G$ of order $n$, if any,
is divisible by $\phi(n)$, the totient of $n$.

Suppose $G$ has one or more elements of order $n$.
Let $N$ be the set $\{x\in G||x|=n\}$.  Then, for any pair
of elements $a,b\in N$, let $a\sim b$ if and only if $a\in\langle b\rangle$.
This defines an equivilance relation on $N$, since $a\in\langle a\rangle$ gives us
the reflexive property, since $a\in\langle b\rangle\implies b\in\langle a\rangle$ gives
us the symmetric property, and since $a\in\langle b\rangle$ and, for $c\in N$, $b\in\langle c\rangle$ implies
that $a\in\langle c\rangle$, giving us the transitive property.  We now note that
by Theorem 4.4, the size of each equivilance class is $\phi(n)$.
It follows that the number of elements of order $n$ is $G$ is
$s\phi(n)$, where $s$ is the number of equivilance classes.

Oh, I had already read this in the book as the Corollary to Theorem 4.4.

\section*{Lemma 4}

If $N$ is a normal subgroup of a group $G$ and
$gN$ for some $g\in G$ is a coset in $G/N$, then
for any $g'\in gN$, we have $gN=g'N$.

If $g'\in gN$, then there exists $n\in N$ such that $g'=gn$.
Then $g'g^{-1}=gng^{-1}\in N$ by the normality of $N$ in $G$,
and it follows that $gN=g'N$ by Property 4 of the Lemma on cosets
in Chapter 7.  Thus any member of a coset can act as a representative
of the coset.

\section*{Exercise 38}

For each pair of positive integers $m$ and $n$, we can define a homomorphism
from $Z$ to $Z_m\oplus Z_n$ by $x\to (x\bmod m, x\bmod n)$.  What is the
kernel when $(m,n)=(3,4)$?  What is the kernel when $(m,n)=(6,4)$?
Generalize.

Let $\phi:Z\to Z_m\oplus Z_n$ be the homomorphism.  Seeing that
\begin{align*}
\ker\phi&=\{x\in Z|\mbox{$x\equiv 0\pmod m$ and $x\equiv 0\pmod n$}\} \\
 &=\{zm|z\in Z\}\cap\{zn|z\in Z\},
\end{align*}
it follows that
\begin{equation*}
\ker\phi=\{z\lcm(m,n)|z\in Z\}.
\end{equation*}

\section*{Exercise 39}

If $K$ is a subgroup of $G$ and $N$ is a normal subgroup of $G$,
prove that $K/(K\cap N)$ is isomorphic to $KN/N$.

Notice that the normality of the subgroup $K\cap N$ in $K$ is proven
by the problem similar to Exercise 50 in Chapter 9.

We now show that $KN$ is a group.
Let $x\in KN$.  Then $x=kn$ for some $k\in K$ and $n\in N$.
But then by Lemma 2 above, $x=n'k\in NK$ for some $n'\in N$.
It follows that $KN\subseteq NK$.
Similarly, we can show that $NK\subseteq KN$, so $NK=KN$.
It then follows by Exercise 6 of the supplementary exercises
for chapters 5 through 8 that $NK$ is a group.

Is $N$ normal in $KN$?

We now let $\phi:K/(K\cap N)\to KN/N$ be a function defined as
\begin{equation*}
\phi(k(K\cap N))=kN,
\end{equation*}
and show that it is a homomorphism.  Let us first verify that this
is a well defined function.  Let $a,b\in K$ such that $a(K\cap N)=b(K\cap N)$.
Then $ab^{-1}\in K\cap N\subseteq N$, showing that $aN=bN$.

We now show that $\phi$ is operation preserving.
By the normality of $N$ and $N\cap K$, we see that
\begin{align*}
 & \phi(a(K\cap N)b(K\cap N)) \\
=\,& \phi(ab(K\cap N)) \\
 =\,& abN=aNbN \\
 =\,& \phi(a(K\cap N))(\phi(b(K\cap N)),
\end{align*}
showing that $\phi$ is operation preserving.

We now consider the kernel of $\phi$.  Notice that
\begin{align*}
\ker\phi &= \{k(K\cap N)\in K/(K\cap N)|k\in N\}, \\
 &= \{k(K\cap N)\in K/(K\cap N)|k\in K\cap N\}, \\
 &= \{K\cap N\}.
\end{align*}
It follows that $\phi$ is an isomorphism by Property 9
of Theorem 10.2.

\section*{Exercise 40}

If $M$ and $N$ are normal subgroups of $G$ and $N\leq M$, prove
that $(G/N)/(M/N)\approx G/M$.

Notice that $M/N$ is a subgroup of $G/N$.  To see that $M/N$ is normal in $G/N$,
let $g\in G$ and let $m\in M$, and see that
\begin{equation*}
gNmN(gN)^{-1}=gmNg^{-1}N=gmg^{-1}N\in M/N,
\end{equation*}
since $gmg^{-1}\in M$ by the normality of $M$ in $G$.

Now consider the mapping $\phi:(G/N)/(M/N)\to G/M$, defined as
\begin{equation*}
\phi(xN(M/N))=yM,
\end{equation*}
where $y$ is any element in the coset $xN$.  Let us now show that
this is a well defined mapping.  Let $a,b\in G$ such that
$aN(M/N)=bN(M/N)$.  It follows that $aN(bN)^{-1}=ab^{-1}N\in M/N\implies ab^{-1}\in M$.
Now let $aN(M/N)$ map to $a'M$ and $bN(M/N)$ map to $b'M$.
Now if $a'\in aN\subseteq aM$, then $a'M=aM$.
Similarly, if $b'\in bN\subseteq bM$, then $b'M=bM$.
But now since $ab^{-1}\in M$, we see that $aM=bM$, so $a'M=b'M$.

Notice that the proof that $\phi$ is well defined also lets us simplify its usage.
That is, for any $x\in G$, we can let $xN(M/N)$ map to $xM$.  This will greatly
ease the remainder of our proof.

We now show that $\phi$ is operation preserving.  Letting $a,b\in G$, we have
\begin{align*}
 & \phi(aN(M/N)bN(M/N)) \\
 =\,& \phi(aNbN(M/N)) \\
 =\,& \phi(abN(M/N)) \\
 =\,& abM = aMbM \\
 =\,& \phi(aN(M/N))\phi(bN(M/N)).
\end{align*}

We now consider the kernel of $\phi$.  We have
\begin{align*}
\ker\phi &=\{gN(M/N)|\mbox{$g\in G$ and $\phi(gN)=M$}\} \\
 &= \{gN(M/N)|g\in M\}.
\end{align*}
Now let $a,b\in M$ and consider $aN(M/N)$ and $bN(M/N)$.
Since $a,b\in M$, we have $ab^{-1}N\in M/N$,
which, in turn, implies that $aN(bN)^{-1}\in M/N\implies aN(M/N)=bN(M/N)$.
It follows that $|\ker\phi|=1$, and therefore, $\phi$ is an isomorphism.

\section*{Exercise 47}

Suppose that for each prime $p$, $Z_p$ is the homomorphic image of a group $G$.
What can we say about $|G|$?  Give an example of such a group.

By Property 6 of Theorem 10.2, we see that $|\phi(G)|$ divides the order of $|G|$.
So, since $\phi(G)=Z_p$, we see that $p$ divides $|G|$.

An automorphism of $Z_p$ may be a trivial example.

After reading the answer in the back of the book, I'm wrong, because I did not understand
the problem statement.  For {\it every} prime $p$, $Z_p$ is {\it a} homomorphic image of {\it the}
group $G$.  So by Property 6 of Theorem 10.2, every prime $p$ divides $|G|$; and since
there are infinitely many primes, $|G|=\infty$.

\section*{Exercise 49}

Let $N$ be a normal subgroup of a group $G$.  Use property 7 of Theorem 10.2
to prove that every subgroup of $G/N$ has the form $H/N$, where $H$ is a subgroup
of $G$.

For every subgroup $H$ of $G$ with $N\leq H$, it is clear that $N$ is normal in $H$
and that $H/N\leq G/N$.  Now let's consider what is somewhat the converse of this.
For every subgroup $K$ of $G/N$, does there exists a subgroup $H$ of $G$
such that $K=H/N$?

Let $\phi:G\to G/N$ be defined as $\phi(g)=gN$.  This is well defined
and operation preserving, so it is a homomorphism from $G$ to $G/N$.
Then, by property 7 of Theorem 10.2, we see that $\phi^{-1}(K)$ is a
subgroup of $G$.  Now notice that if $n\in N$, then $\phi(n)=nN=N\in K$,
showing that $N\leq\phi^{-1}(K)$.  It follows that $N$ is normal in $\phi^{-1}(K)$.
Letting $H=\phi^{-1}(K)$, what remains to be shown now is that $H/N=K$.  Letting $g\in G$, we have
\begin{equation*}
gN\in\phi^{-1}(K)/N\iff g\in\phi^{-1}(K)\iff\phi(g)\in K\iff gN\in K.
\end{equation*}
It follows that $H/N=K$.

\section*{Exercise 50}

Show that a homomorphism defined on a cyclic group is completely determined by its action on a generator of the group.

Let $\phi:G\to\overline{G}$ be a homomorphism of the cyclic group $G=\langle g\rangle$.
Then $\phi(g^k)=\phi(g)^k$.

\section*{Exercise 51}

Use the First Isomorphism Theorem to prove Theorem 9.4.

We want to show that for any group $G$, $G/Z(G)\approx\inn(G)$.
To that end, let $\phi_g(x)=gxg^{-1}\in\inn(G)$, and define $\psi(g)=\phi_g$.
It is clear that $\psi$ is a homomorphism from $G$ onto $\inn(G)$.
Now realize that $\psi(g)=\phi_e$ if and only if $gxg^{-1}=e$,
showing that $\ker\psi=Z(G)$.  It now follows from Theorem 9.4
that $G/G(Z)=G/\ker\psi\approx\psi(G)=\inn(G)$.

\section*{Exercise 52}

Let $\alpha$ and $\beta$ be group homomorphisms from $G$ to $\overline{G}$ and let
$H=\{g\in G|\alpha(g)=\beta(g)\}$.  Prove or disprove that $H$ is a subgroup of $G$.

Clearly $e\in H$ by Property 1 of Theorem 10.1.
Now let $a,b\in H$.  We then have
\begin{equation*}
\alpha(ab^{-1})=\alpha(a)\alpha(b)^{-1}=\beta(a)\beta(b)^{-1}=\beta(ab^{-1}),
\end{equation*}
showing that $ab^{-1}\in H$.  So I think it's a subgroup of $G$.

\section*{Exercise 54}

If $H$ and $K$ are normal subgroups of $G$ and $H\cap K=\{e\}$, prove that $G$
is isomorphic to a subgroup of $G/H\oplus G/K$.

Let $\phi(g)=(gH,gK)$.  It is clear that $\phi$ is a homomorphism from $G$
to $G/H\oplus G/K$.  Then, by property 1 of Theorem 10.2, we see that
$\phi(G)$ is a subgroup of $G/H\oplus G/K$.  We will now show that $G\approx\phi(G)$,
and do so by showing that $\phi$ is an isomorphism between $G$ and $\phi(G)$.
Notice that $\phi$ is clearly onto.  We already know it is operation preserving.
All that remains to be shown then is that $\phi$ is one-to-one.  So, let $a,b\in G$ such that
$(aH,aK)=(bH,bK)$.  Then $aH=bH$ and $aK=bK$, so $ab^{-1}\in H$ and $ab^{-1}\in K$.
It follows that $ab^{-1}\in H\cap K\implies ab^{-1}=e\implies a=b$, showing that
$\phi$ is one-to-one.

\section*{Exercise 55}

Suppose that $H$ and $K$ are distinct subgroups of $G$ of index 2.  Prove that
$H\cap K$ is a normal subgroup of $G$ of index 4 and that $G/(H\cap K)$ is not cyclic.

It is easy to show that both $H$ and $K$ are normal in $G$.  By symmetry of the
problem, we need only show that $H$ is normal in $G$.  (The proof for $K$ is similar.)
Let $g\in G-H$.  Then $gH\neq H$ and $Hg\neq H$.  But $gH$ is the only remaining
of the 2 cosets of $H$ in $G$, so $gH=Hg$, and therefore $H$ is normal in $G$.

We then see that $H\cap K$ is normal in $G$ by Exercise 50 of chapter 9.

At this point I had to peak at the back of the book to get some hints.
By Exercise 39 in this chapter, we see that $K/(K\cap H)\approx KH/H$,
so $|K|/|K\cap H|=|KH|/|H|$.  But $|H|=|K|=|G|/2$, so we see that
$|G|^2/4=|KH||K\cap H|$.  Now realize that $|KH|=|G|$, because $K\neq H\implies |HK|>|G|/2$,
and $KH=HK$ implies that $HK$ is a subgroup of $G$, and so $|HK|$ divides $|G|$.

Lastly, the back of the book says that $G/(H\cap K)$ has two subgroups of order 2,
which clearly means that it can't by cyclic by the Fundamental Theorem of Cyclic Groups.
To which two subgroups of $G/(H\cap K)$ is Gallian refering?  I'm not entirely sure,
and need to go to bed.

\end{document}