\documentclass{birkjour}

\usepackage{tikz}
\usepackage{graphicx}
\usepackage{hyperref}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\newtheorem*{ex}{Example}
\numberwithin{equation}{section}

\newcommand{\R}{\mathbb{R}}
\newcommand{\B}{\mathbb{B}}
\newcommand{\G}{\mathbb{G}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\gd}{\dot{g}}
\newcommand{\gh}{\hat{g}}
\newcommand{\Gd}{\dot{G}}
\newcommand{\Gh}{\hat{G}}
\newcommand{\nvai}{\infty}
\newcommand{\nvao}{o}
\newcommand{\grade}{\mbox{grade}}

\begin{document}

%-------------------------------------------------------------------------
% editorial commands: to be inserted by the editorial office
%
%\firstpage{1} \volume{228} \Copyrightyear{2004} \DOI{003-0001}
%
%
%\seriesextra{Just an add-on}
%\seriesextraline{This is the Concrete Title of this Book\br H.E. R and S.T.C. W, Eds.}
%
% for journals:
%
%\firstpage{1}
%\issuenumber{1}
%\Volumeandyear{1 (2004)}
%\Copyrightyear{2004}
%\DOI{003-xxxx-y}
%\Signet
%\commby{inhouse}
%\submitted{March 14, 2003}
%\received{March 16, 2000}
%\revised{June 1, 2000}
%\accepted{July 22, 2000}
%
%
%
%---------------------------------------------------------------------------

\title{The Intersection Of Rays And\\Algebraic Surfaces}

\author{Spencer T. Parkin}
\address{102 W. 500 S., Salt Lake City, UT  84101}
\email{spencerparkin@outlook.com}

\subjclass{Primary 14J70; Secondary 14J29}

%\date{October 22, 2013}

%\dedicatory{To my dear wife Melinda.}

\begin{abstract}
It is shown that for any multi-variable polynomial defined over the
real numbers that the image of a line through the domain of such a function
is determined entirely by all orders of the directional derivatives of this function at
any one point along the line and in a direction of the line.  This result has an
application in the problem of casting rays through algebraic surfaces as it
shows that such a problem, in all cases, reduces to the problem of finding
the roots of a single-variable polynomial having an explicit formulation
in terms of the multi-variable polynomial and ray in question.
\end{abstract}

\maketitle

% More needs to be said at start, especially about the notation so that
% sub-scripts are understood.

\section{Introduction}

Letting $f:\R^n\to\R$ be any polynomial equation in $n$ variables and up to degree $m$, it
was shown in \cite{Parkin13} that the function $\bigwedge_{i=1}^m p_i(x)$ may be factored out
of this polynomial in terms of the inner product as
\begin{equation}\label{equ_polynomial_func}
f(x) = \bigwedge_{i=1}^m p_i(x)\cdot B,
\end{equation}
where $B$ is an $m$-vector of our geometric algebra with $\nvai_i\cdot B=0$ for all
integers $i\in[1,m]$, and where the function
$p_i:\R^n\to\V_i$ is given by
\begin{equation}\label{equ_point_func}
p_i(x) = \nvao_i + x_i + \frac{1}{2}x_i^2\nvai_i,
\end{equation}
having its origins in the paper \cite{Hestenes01}.
Given a point $x\in\R^n$ and a direction vector $v\in\R^n$,
we wish to find the set of all scalars $\lambda\in\R$ such that $f(x+\lambda v)=0$.
Utilizing equation \eqref{equ_polynomial_func} for this purpose, we easily find that
\begin{equation}\label{equ_ray_polynomial_unexpanded}
f(x+\lambda v) = \bigwedge_{i=1}^m(p_i(x)+\lambda v_i)\cdot B,
\end{equation}
because we can ignore the $\frac{1}{2}x_i^2\nvai_i$ term in equation \eqref{equ_point_func}.
Looking at equation \eqref{equ_ray_polynomial_unexpanded},
it is immediately clear that its expansion is that of a polynomial in $\lambda$ of up to degree $m$.
What we're going to show in this paper is that an explicit formula for this polynomial can
be found in terms of all orders of directional derivatives of $f$ at $x$ and in the
direction of $v$.

\section{The Result}

We begin by rewriting equation \eqref{equ_ray_polynomial_unexpanded} as
\begin{equation}\label{equ_ray_polynomial_expanded_but_unknown}
f(x+\lambda v) = \sum_{i=0}^m T_i(x),
\end{equation}
where $T_i(x)$ will denote the $i^{th}$ term involving $\lambda^i$ in the series expansion of
\eqref{equ_ray_polynomial_unexpanded}.  Carefully formulating this term, we get
\begin{equation*}
T_i(x) = \lambda^i \sum_{j=1}^{\binom{m}{i}} W_{j,i}(x)\cdot B,
\end{equation*}
where $W_{j,i}$ is the $j^{th}$ way to write an outer product involving
$i$ vectors taken from $\{v_k\}_{k=1}^m$ and $m-i$ vectors taken from
$\{p_k(x)\}_{k=1}^m$ in an order having ascending sub-scripts.  The following
examples help clarify this in the case $m=3$.
\begin{align*}
W_{1,0} &= p_1\wedge p_2\wedge p_3 \\
W_{1,1} &= p_1\wedge p_2\wedge v_3 \\
W_{2,1} &= p_1\wedge v_2\wedge p_3 \\
W_{3,1} &= v_1\wedge p_2\wedge p_3 \\
W_{1,2} &= p_1\wedge v_2\wedge v_3 \\
W_{2,2} &= v_1\wedge p_2\wedge v_3 \\
W_{3,2} &= v_1\wedge v_2\wedge v_3 \\
W_{1,3} &= v_1\wedge v_2\wedge v_3
\end{align*}
Having now come to terms, (no pun intended), with the general expansion of
equation \eqref{equ_ray_polynomial_unexpanded}, we proceed now to
fearlessly take the directional derivative of $T_i$ at $x$ and in the direction of $v$.
Doing so, we get
\begin{align*}
\nabla_v T_i(x) &= \lambda^i\sum_{j=1}^{\binom{m}{i}}\lim_{\delta\to 0}
\frac{W_{j,i}(x+\delta v)-W_{j,i}(x)}{\delta}\cdot B,
\end{align*}
knowing that each individual limit will exist.  What we must realize now is that
the term $W_{j,i}(x)$ will get canceled in the expansion
of $W_{j,i}(x+\delta v)$, leaving only terms that are multiples of positive powers of $\delta$.
Furthermore, it is only those remaining terms that are multiples of $\delta$ itself
that will survive the limit process.  We are therefore left to deduce these terms in an
evaluation of the limit.  What we find is that all such terms are of the form
$\delta W_{j,i+1}(x)$, but we need to determine just how many we have.
Realizing that $\binom{m}{i}$ old terms will each contribute $m-i$ new terms of
this form, of which there should be $\binom{m}{i+1}$, but that no type of term
will be produced any more or less than any other, we see that
\begin{equation*}
\frac{(m-i)\binom{m}{i}}{\binom{m}{i+1}}=i+1
\end{equation*}
is the number of such terms of the form $\delta W_{j,i+1}(x)$, and we may write
\begin{align}
\nabla_v T_i(x) &= \lambda^i\sum_{j=1}^{\binom{m}{i+1}}\lim_{\delta\to 0}
\frac{(i+1)\delta W_{j,i+1}(x)}{\delta}\cdot B\nonumber \\
 &= \lambda^i(i+1)\sum_{j=1}^{\binom{m}{i+1}} W_{j,i+1}(x)\cdot B\nonumber \\
 &= \frac{i+1}{\lambda}T_{i+1}(x).\label{equ_recurrence_relation}
\end{align}
Returning to equation \eqref{equ_ray_polynomial_expanded_but_unknown},
and realizing that $T_0(x)=f(x)$, we can now finally deduce the expansion of
\eqref{equ_ray_polynomial_unexpanded} using the recurrence relation
of equation \eqref{equ_recurrence_relation} as
\begin{equation}\label{equ_result}
f(x+\lambda v) = \sum_{i=0}^m\frac{\lambda^i}{i!}\nabla_v^i f(x),
\end{equation}
where $\nabla_v^i f(x)$ is the $i^{th}$ order directional derivative of $f$ at $x$
in the direction of $v$ with $\nabla_v^0 f(x)=f(x)$.

\section{Making Use Of The Result}

In its present form, equation \eqref{equ_result} lacks ease of use, because knowledge of the
vector $v$ is needed to calculate all orders of the directional derivative.  To solve this problem,
we need to decouple the knowledge of this vector from the limit processes by generalizing the
idea of the gradient to higher orders.  This has already been done in \cite{} with the use of
$k$-forms.  The manifestation of such things in the algebra we are using may be as follows.

We begin with the gradient of $f$, usually written $\nabla f$.  But this is ambiguous in our algebra,
because we need to specify the sub-algebra over which $\nabla f$ will be taken.  We will
do this with an integral subscript as $\nabla_i f$, not to be confused with the directional
derivative, which uses a vector subscript.  The gradient can now be defined similarly to
it's definition in \cite{} as
\begin{equation*}
\nabla_i f(x) = \sum_{j=1}^n e_{i,j}\nabla_{e_j} f(x).
\end{equation*}
Having done this, we have, for any integer $i\in[1,m]$,
\begin{equation*}
\nabla_v f(x) = v_i\cdot\nabla_i f(x),
\end{equation*}
which is a well-known result.  Generalizing this, we find that
\begin{equation*}
\nabla_v^i f(x) = \bigwedge_{j=1}^i v_j\cdot\left(\bigwedge_{j=1}^i\nabla_j\right) f(x),
\end{equation*}
where the function operator $\bigwedge_{j=1}^i\nabla_j$ is defined as
\begin{equation*}
\left(\bigwedge_{j=1}^i\nabla_j\right)f(x) =
-(-1)^i \sum_{j_1=1}^n\dots\sum_{j_i=1}^n\bigwedge_{k=1}^i e_{k,j_k}\left(\prod_{k=1}^i\nabla_{e_{j_k}}\right)f(x),
\end{equation*}
realizing that, as was assumed throughout \cite{}, the outer and inner products bind tighter than
the geometric product.
Here, the function operator being applied to $f$ is an $i^{th}$ order partial derivative.

The decoupling of equation \eqref{equ_result} can now be written as
\begin{equation}\label{equ_result_decoupled}
f(x+\lambda v) = \sum_{i=0}^m\frac{\lambda^i}{i!}\bigwedge_{j=1}^i v_j\cdot\left(\bigwedge_{j=1}^i\nabla_j\right)f(x),
\end{equation}
where the outer product $\bigwedge_{j=1}^i v_j$ is one in the case $i=0$, and where the function
operator $\bigwedge_{j=1}^i\nabla j$ is the identity operator in the same case.

If all orders of the gradient of $f$ are available to us, equation \eqref{equ_result_decoupled} becomes
a convenient way to calculate the coeficients of the polynomial whose zeros give us the parameters of
the intersection points of our ray with the a given algebraic surface.

Looking back again at equation \eqref{equ_ray_polynomial_unexpanded}, this is certainly a usable form
if a symbolic calculator is available.  Use of symbolic calculation was employed in \cite{}.  If no such
thing is available, then \eqref{equ_result_decoupled} may allow us to come up with the polynomial
through literal evaluation, provided, again, that we have all gradients available to us.
For algebraic surfaces of degrees greater than or equal to five\footnote{There is no closed form solution
to a polynomial of degree greater than or equal to five in terms of elementary functions.  See\cite{}.}, we may
lose the need to come up
with the polynomial expansion of $f(x+\lambda v)$ altogether in favor of root-finding methods
that need know nothing more about $f(x+\lambda v)$ than that it is continuous.\footnote{In such a
case, the intermediate value theorem applies.  Recall that the composition of two continuous functions
is continuous.}

In any case, \eqref{equ_result_decoupled} appears to be, if nothing more, an
interesting result about algebraic surface that may find applications in other areas.

\bibliographystyle{amsplain}
\bibliography{Parkin_TheIntersectionOfRaysAndAlgebraicSurfaces}

\end{document}