\documentclass{birkjour}

\usepackage{tikz}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{algpseudocode}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\newtheorem*{ex}{Example}
\numberwithin{equation}{section}

\newcommand{\R}{\mathbb{R}}
\newcommand{\B}{\mathbb{B}}
\newcommand{\G}{\mathbb{G}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\gd}{\dot{g}}
\newcommand{\gh}{\hat{g}}
\newcommand{\Gd}{\dot{G}}
\newcommand{\Gh}{\hat{G}}
\newcommand{\nvai}{\infty}
\newcommand{\nvao}{o}
\newcommand{\grade}{\mbox{grade}}

%\received{}\accepted{}

\begin{document}

\title{An Introduction To Geometric Sets}

\author{Spencer T. Parkin}
\address{102 W. 500 S., \\
Salt Lake City, UT  84101} \email{spencerparkin@outlook.com}

%\subjclass{Primary 14J70; Secondary 14J29}

%\dedicatory{To my dear wife Melinda.}

\begin{abstract}
A general theory of geometric models based upon the idea of using blades of a geometric algebra
as representatives of geometry is developed.  Similar to the idea of an algebraic set, the main object of study
becomes the notion of a geometric set.  Results about geometric sets are obtained, and their implications discussed.
\end{abstract}

\keywords{Geometric Algebra, Model Of Geometry, Geometric Set}

\maketitle

\section{Introduction}

This paper develops a general theory of geometric models
based upon the idea of using blades of a geometric algebra as representatives of geometry.
Specific examples include the homogeneous and conformal models.  (See \cite{}.)

\subsection{Motivation}

Seeing a great deal of commonality between various models of geometry based upon
geometric algebra, a formal treatment of the subject deserves to be given in an abstract
setting in much the same way that, for example, abstract algebra provides such a setting
in which algebraic sets generated by ideals of a polynomial ring can be studied.
Readers familiar with algebraic geometry will no-doubt recognize at least some small overlap
between that subject and the subject of this paper.  For example, it will be seen
that in some cases, geometric sets are algebraic sets.

\subsection{Conventions}

This paper uses capital letters $A,B,C$ to denote blades of various grades,
while using lower case letters $a,b,c$ to denote vectors.  Scalars are written
using greek letters $\alpha,\beta,\gamma$.  Grades of blades $A,B,C$ are usually
denoted by lower case letters $r,s,t$, respectively, unlesss stated otherwise.  Lower case letters $i,j,k$ are
used as indices.  Capital letters $R,S,T$
are used to denote subsets of interest of $n$-dimensional euclidean space $\R^n$.
Lower case letters $x,y,z$ are reserved for denoting points taken from $\R^n$.  We
let $P(\R^n)$ denote the power set of $\R^n$.
We will use $\G$ to denote our geometric algebra, and $\V$ to denote an $m$-dimensional vector
space generating it.  $\B$ will denote the set of all blades taken from $\G$.  The
scalars of $\V$, and therefore $\G$, are taken from the field of real numbers $\R$.
The capital letter $I$ will be used to denote the unit psuedo-scalar of $\G$.  We assume
that $I$ is invertible with respect to the geometric product.

We will let the geometric product take precedence over the inner and outer products,
and the inner product take precedence over the outer product.

No specific signature of our geometric algebra $\G$ is assumed in this paper unless
one is given in a special case.

At times a blade $A$ may be referred to as a subspace.  The notion of blades as being
representatives of subspaces of $\V$ is a common practice and is employed throughout this paper.
It may also be said that a blade $A$ is a subspace of some other blade $B$; in which case,
it can be understood that for all vectors $a\in\V$ such that $a\wedge A=0$, we have $a\wedge B=0$.

\section{Preliminaries}

The results of this paper will depend upon the following preliminary material.
If desired, the reader is welcome to skip this material and refer back to it only as needed.

\subsection{Identities}

Given a vector $a\in\V$ and a blade $A\in\B$, central to all of geometric algebra is the identity
\begin{equation}\label{equ_vec_blade_geo_prod}
aA = a\cdot A + a\wedge A.
\end{equation}
The inner and outer products of \eqref{equ_vec_blade_geo_prod} may be written in terms
of the geometric product as
\begin{align}
a\wedge A &= \frac{1}{2}\left(aA + (-1)^rAa\right),\label{equ_vec_blade_outer_prod} \\
a\cdot A &= \frac{1}{2}\left(aA - (-1)^rAa\right),\label{equ_vec_blade_inner_prod}
\end{align}
where $r=\grade(A)$.  Then, realizing that $m=\grade(I)$, and that by \eqref{equ_vec_blade_geo_prod},
we have $aI=a\cdot I$, we can use equation \eqref{equ_vec_blade_inner_prod} to establish the
commutativity of vectors in $\V$ with the unit psuedo-scalar $I$ as
\begin{equation}\label{equ_blade_psuedo_scalar_commutativity}
aI = -(-1)^mIa.
\end{equation}
Using equation \eqref{equ_blade_psuedo_scalar_commutativity} in conjunction with equation \eqref{equ_vec_blade_inner_prod},
we find that
\begin{equation}\label{equ_vec_blade_inner_prod_dual}
(a\cdot A)I=a\wedge AI.
\end{equation}
(In verifying this identity, it helps to realize that for any integer $r$, we have $(-1)^r=(-1)^{-r}$.)
Replacing $A$ in equation \eqref{equ_vec_blade_inner_prod_dual} with $AI$, we find that
\begin{equation}\label{equ_vec_blade_outer_prod_dual}
(a\wedge A)I=a\cdot AI.
\end{equation}

Refering back to equation \eqref{equ_vec_blade_inner_prod}, another important formulation of the inner product between a vector and a blade
is given by
\begin{equation}\label{equ_vec_blade_inner_prod_series}
a\cdot A=-\sum_{i=1}^r(-1)^i(a\cdot a_i)A_i,
\end{equation}
where $A$ is factored as $\bigwedge_{i=1}^r a_i$, and we define $A_i$ as
\begin{equation}\label{equ_A_i}
A_i=\bigwedge_{\substack{j=1\\j\neq i}}^r a_i.
\end{equation}
This leads to the following recursive formulation.
\begin{equation}
a\cdot A = (a\cdot a_1)A_1-a_1\wedge(a\cdot A_1)
\end{equation}
If a blade $B\in\B$ has grade $s$ and factorization $\bigwedge_{i=1}^s b_i$, then we can express
the product $A\cdot B$ recursively as
\begin{equation}
A\cdot B = \left\{\begin{array}{ll}
A_r\cdot(a_r\cdot B) & \mbox{if $r\leq s$,} \\
(A\cdot b_1)\cdot B_1 & \mbox{if $r\geq s$.}
\end{array}\right.
\end{equation}

\subsection{Lemmas}

The following lemmas will help us prove the results of this paper.

\begin{lem}[Found Factorization Of Blade]\label{lem_desired_blade_factorization}
If $A\in\B$ is a non-zero blade of grade $r>0$ and $\{c_i\}_{i=1}^r$ is a set
of $r$ linearly independent vectors such that for all $c\in\{c_i\}_{i=1}^r$, we have
$c\wedge A=0$, then there exists a scalar $\beta\in\R$ such that $A=\beta C$,
where $C$ is an $r$-blade given by
\begin{equation}\label{equ_desired_blade_factorization}
C=\bigwedge_{i=1}^r c_i.
\end{equation}
\end{lem}
\begin{proof}
Letting $\bigwedge_{i=1}^r a_i$ be a factorization of the $r$-blade $A$, it is clear
that if $c_i\wedge A=0$, then there exists a set of $r$ scalars $\{\gamma_{i,j}\}_{j=1}^r$
such that
\begin{equation}\label{equ_c_i}
c_i=\sum_{j=1}^r\gamma_{i,j}a_i.
\end{equation}
We then have
\begin{equation}
\bigwedge_{i=1}^r c_i=(\det M)\bigwedge_{i=1}^r a_i,
\end{equation}
where the $r\times r$ matrix $M$ is given by
\begin{equation}
M = \left[\begin{array}{cccc}
\gamma_{1,1} & \gamma_{1,2} & \dots & \gamma_{1,r} \\
\gamma_{2,1} & \gamma_{2,2} & \dots & \gamma_{2,r} \\
\vdots & \vdots & \ddots & \vdots \\
\gamma_{r,1} & \gamma_{r,2} & \dots & \gamma_{r,r}
\end{array}\right].
\end{equation}
Now since $\{c_i\}_{i=1}^r$ is a linearly independent set of vectors,
$\det M\neq 0$, and we may choose $\beta=(\det M)^{-1}$ and
equation \eqref{equ_desired_blade_factorization} now holds.
\end{proof}

Our next lemma is a generalization of Lemma~\ref{lem_desired_blade_factorization}.

\begin{lem}[Found Partial/Full Factorization Of Blade]\label{lem_desired_partial_blade_factorization}
If $A\in\B$ is a non-zero blade of grade $r>0$ and $\{c_i\}_{i=1}^t$ is a set of $t\leq r$ linearly
independent vectors such that for all $c\in\{c_i\}_{i=1}^t$, we have $c\wedge A=0$, then there
exists a blade $B\in\B$ of grade $r-t$ such that $A=B\wedge C$, where $C$ is a $t$-blade given by
\begin{equation}
C = \bigwedge_{i=1}^t c_i.
\end{equation}
\end{lem}
\begin{proof}
The case of $t=r$ being handled by Lemma~\ref{lem_desired_blade_factorization}, and the
case of $r=1$ being trivial, we may assume that $0<t<r$.  Then, letting $\bigwedge_{i=1}^r a_i$
be a factorization of the $r$-blade $A$, it is clear that if $c_i\wedge A=0$, then there exists a
set of $r$ scalars $\{\gamma_{i,j}\}_{j=1}^r$ such that equation \eqref{equ_c_i} again holds.
Furthermore, we can assume, without loss of generality, that for each $c_i$, $\gamma_{i,i}\neq 0$.
We then have
\begin{equation}
\bigwedge_{i=1}^t c_i = \gamma\bigwedge_{i=1}^t a_i + Q,
\end{equation}
where $\gamma=\prod_{i=1}^t\gamma_{i,i}$ and $Q$ represents the remaining terms in the
expansion of $C$.  Letting
\begin{equation}
B=(-1)^{t(r-t)}\gamma^{-1}\bigwedge_{i=t+1}^r a_i,
\end{equation}
it follows now that $A=B\wedge C$, since $B\wedge Q=0$.
\end{proof}

% may want to note that 0-blades in outer, inner and geometric products reduce to scalar-blade product

\begin{lem}\label{lem_lin_indep_blades}
For a non-zero $r$-blade $A$ factored as $\bigwedge_{i=1}^r a_i$, the set of
$(r-1)$-blades $\{A_i\}_{i=1}^r$, (see equation \eqref{equ_A_i}), is linearly independent.
\end{lem}
\begin{proof}
Suppose there exists a non-trivial set of $r$ scalars $\{\alpha_i\}_{i=1}^r$
such that $0 = \sum_{i=1}^r\alpha_i A_i$.
Then, without loss of generality, suppose that $\alpha_r\neq 0$, and rearrange
our equation as $-\alpha_rA_r=\sum_{i=1}^{r-1}\alpha_iA_i$.  Now notice
that while $a_r\wedge-\alpha_rA_r\neq 0$, we have $a_r\wedge\sum_{i=1}^{r-1}\alpha_iA_i=0$,
which is a contradiction.
\end{proof}

Note that a perhaps more elegant proof of Lemma~\ref{lem_lin_indep_blades} could have been given
under the assumption
of a euclidean signature; in which case, the Gram-Schmidt orthogonalization process would have
allowed us to choose, without loss of generality, an orthogonal factorization of the blade $A$.
Doing so, $A$ becomes the versor $\prod_{i=1}^r a_i$, and we may write
\begin{equation}\label{equ_elegant_proof_of_A_i_lin_indep}
0 = \sum_{i=1}^r \alpha_iA_i \iff
0=\left(\sum_{i=1}^r\alpha_iA_i\right)A = -\sum_{i=1}^r(-1)^{r-i}A_i^2\alpha_ia_i.
\end{equation}
The use of equation \eqref{equ_elegant_proof_of_A_i_lin_indep} would depend, in part, upon Lemma~\ref{lem_zero_prod_property} to follow.

\begin{lem}[The Zero Product Property]\label{lem_zero_prod_property}
For any two non-zero blades $A,B\in\B$ of grades $r$ and $s$, respectively, if $AB=0$ and at least one of $A$ and $B$
is invertible, then at least one of $A$ and $B$ is zero.
\end{lem}
\begin{proof}
Without loss of generality, suppose $B^{-1}$ exists.  We then see that
\begin{equation}
A=ABB^{-1}=0B^{-1}=0.
\end{equation}
\end{proof}

Notice the requirment here of Lemma~\ref{lem_zero_prod_property} that at least one of $A$ and $B$ be invertible.
This requirement comes about in consideration of the square of a non-zero null-vector.

\section{Results}

\subsection{Foundation}

Our discussion begins with a non-zero, undefined function $p:\R^n\to\V$.\footnote{By definition, a function is well-defined even if it is left unspecified.}
By leaving this function undefined, the results to follow generalize to the homogeneous model, conformal model,
and any other model of geometry that is based upon the use of blades as representatives of geometry.

\begin{defn}[Direct And Dual Representation]\label{def_gh_and_gd}
For the two functions $\gh:\B\to P(\R^n)$ and $\gd:\B\to P(\R^n)$, given by
\begin{align}
\gh(A) &= \{x\in\R^n|p(x)\wedge A=0\},\\
\gd(A) &= \{x\in\R^n|p(x)\cdot A=0\},
\end{align}
we say that $A$ directly represents the set of points $\gh(A)$ and
dually represents the set of points $\gd(A)$.
\end{defn}

From Definition~\ref{def_gh_and_gd}, it is important to take away the realization that a given blade $A\in\B$
represents two subsets of $\R^n$ simultaneously; namely, $\gh(A)$ and $\gd(A)$.  Which we choose
to think of $A$ as being a representative of at any given time is completely arbitrary.

It should also be clear from Definition~\ref{def_gh_and_gd} that the subset of $\R^n$ represented by a
blade $A$, (directly or dually), remains invariant under any non-zero scaling of the blade $A$.

Finally, now enters this paper's object of study: the {\it geometric set}.

\begin{defn}[Geometric Set]\label{def_geo_set}
A subset $R\subset\R^n$ for which there exists a blade $A\in\B$ such that $\gh(A)=R$ is what
we'll refer to as a {\it geometric set}.
\end{defn}

In the course of our study, we will find the concept of {\it irreducibility} important.

\begin{defn}[Irreducible/Reducible Blade]\label{def_irreducible_blade}
Given an $r$-blade $A\in\B$, if there does not exist an $s$-blade $B\in\B$ with $s<r$
such that $\gh(A)=\gh(B)$, then $A$ is what we'll refer to as an {\it irreducible} blade.
A blade that is not irreducible is referred to as {\it reducible}.
\end{defn}

\subsection{Developments}

Having given the definitions in the previous section, we may now focus on the results that follow
from these definitions.

\subsubsection{Representation}

Our initial developments reveal results about the representations of geometric sets.

\begin{lem}[Dual Relationship Between Representations]\label{lem_dual_relationship_in_rep}
For any geometric set $R\subset\R^n$, if $A\in\B$ is a blade such that $\gh(A)=R$,
then $\gd(AI)=R$, and similarly, if $A\in\B$ is a blade such that $\gd(A)=R$, then $\gh(AI)=R$.
\end{lem}
\begin{proof}
By the identity of equation \eqref{equ_vec_blade_outer_prod_dual}, and Lemma~\ref{lem_zero_prod_property},
the first of these two latter statements is proven by
\begin{equation}
0=p(x)\wedge A=-(p(x)\cdot AI)I\iff p(x)\cdot AI=0,
\end{equation}
while the second, by the identity of equation \eqref{equ_vec_blade_inner_prod_dual},
and again Lemma~\ref{lem_zero_prod_property}, is proven by
\begin{equation}
p(x)\cdot A=0\iff 0=(p(x)\cdot A)I=p(x)\wedge AI.
\end{equation}
\end{proof}

In other words, Lemma~\ref{lem_dual_relationship_in_rep} is telling us that for a single given geometric set, the
algebraic relationship between a blade directly (dually) representative of that set,
and a blade dually (directly) representative of that set, is simply that, up to scale, they
are duals of one another.

Of course, there will also be a geometric relationship between the geometric set that is directly represented by a single
given blade $A\in\B$, and the geometric set that is dually represented by $A$, but this would depend upon how we choose to define
the function $p:\R^n\to\V$.

\begin{lem}
For any geometric set $R\subset\R^n$, there exists a blade $A\in\B$ such that
$\gd(A)=R$.
\end{lem}
\begin{proof}
Letting $B\in\B$ be a blade such that $\gh(B)=R$, simply let $A=BI$, and our
lemma goes through by Lemma~\ref{lem_dual_relationship_in_rep}.
\end{proof}

Our next lemma shows that there is no overlap between the geometric sets dually and directly
represented by a blade.

\begin{lem}
For all invertible blades $A\in\B$, we have
\begin{equation}
\gh(A)\cap\gd(A)=\emptyset.
\end{equation}
\end{lem}
\begin{proof}
Supposing $x\in\gh(A)\cap\gd(A)$, we see that
\begin{equation}
0 = p(x)\cdot A + p(x)\wedge A = p(x)A,
\end{equation}
but $p(x)$ is non-zero and $A$ is invertible and therefore non-zero.  We therefore reach a
contradiction by Lemma~\ref{lem_zero_prod_property}.
\end{proof}

\begin{lem}[The Point-Fitting Lemma]\label{lem_factorization_of_irreducible_blades}
If $A\in\B$ is an irreducible $r$-blade with $\gh(A)\neq\emptyset$, then there exists a set of $r$ points $\{x_i\}_{i=1}^r\subset\R^n$
and a scalar $\beta\in\R$ such that
\begin{equation}
A=\beta\bigwedge_{i=1}^r p(x_i).
\end{equation}
\end{lem}
\begin{proof}
Let $t$ be the largest integer for which there exists a set of $t$ points $\{x_i\}_{i=1}^t\subset\gh(A)$ such that
$\bigwedge_{i=1}^t p(x_i)\neq 0$.
Clearly $t\geq 1$, because $\gh(A)$ is non-empty; and clearly $t\leq r$ because of
the requirement that $\{p(x_i)\}_{i=1}^t$ be a linearly independent set with each $p(x_i)\wedge A=0$.  Now if $t=r$, we're
done by Lemma~\ref{lem_desired_blade_factorization}.  Therefore,
supposing $t<r$, there must exist, by Lemma~\ref{lem_desired_partial_blade_factorization}, a factorization of $A$ of the form
$A=B\wedge C$, where $B$ is a blade of grade $r-t$, and $C$ is a $t$-blade given by
\begin{equation}
C=\bigwedge_{i=1}^t p(x_i),
\end{equation}
Now realize that $\gh(A)\subseteq\gh(C)$ or else $t$ is not the largest of its kind\footnote{Suppose
there exists $x\in\gh(A)$ with $x\not\in\gh(C)$.  Then $p(x)\wedge C\neq 0$ and we have found $t+1$
points, (namely, those in $\{x\}\cup\{x_i\}_{i=1}^t$), for which the set of vectors $\{p(x)\}\cup\{p(x_i)\}_{i=1}^t$ is a linearly independent set.}, and
that $\gh(C)\subseteq\gh(A)$, because $C$ is a subspace of $A$.
It now follows that $\gh(A)=\gh(C)$, which contradicts the irreducibility of the blade $A$.
\end{proof}

Notice that in the proof of Lemma~\ref{lem_factorization_of_irreducible_blades} that $\gh(B)=\emptyset$.  It is important to realize, however,
that although there is no point $x\in\gh(A)$ such that $x\in\gh(B)$, this does not imply that $x\in\gh(C)$.
To see why, consider equation \eqref{equ_geo_set_union} below.\footnote{Put another way,
realize that while $(e_1+e_2)\wedge e_1\wedge e_2=0$, the vector $e_1+e_2$ is not in the subspace
spanned by $e_1$, nor that of $e_2$.}

Interestingly, Lemma~\ref{lem_factorization_of_irreducible_blades} shows us that every geometric
set is determined by a finite subset of its points.  For any given definition of $p:\R^n\to\V$, this
would make knowing, for any given set of points $\{x_i\}$, when $\bigwedge_i p(x_i)\neq 0$,
an important result to establish.

It was noted earlier that if $A,B\in\B$ are blades such that for a scalar $\beta\in\R$,
we have $A=\beta B$, then $\gh(A)=\gh(B)$.  The converse of this statement, however,
is not generally true, but leads us to an important and fundamental theorem.

\begin{thm}[The Fundamental Theorem Of Geometric Set Representation]\label{thm_geo_set_rep}
For every geometric set $R\subset\R^n$, there exists, up to scale, a unique,
irreducible blade $A\in\B$ such that $\gh(A)=R$.  Moreover, this blade $A$ is a
subspace of every blade $B$ directly representative of $R$.
\end{thm}
\begin{proof}
By the proof of Lemma~\ref{lem_factorization_of_irreducible_blades}, it is not hard to see that, given any blade $B\in\B$ directly
representative of the geometric set $R$, a subspace $A$ of $B$ can be found that is
also representative of $R$ while having the property of being irreducible.

Suppose now that $A$ and $A'$ are two independently found blades directly representative of $R$
and each irreducible.  It can be easily understood that $\grade(A)=\grade(A')$ by Defintion~\ref{def_irreducible_blade}.
By Lemma~\ref{lem_factorization_of_irreducible_blades}, the $r$-blade $A$ has a factorization
of the form $\alpha\bigwedge_{i=1}^r p(x_i)$.  But now for each $x_i$, we have $p(x_i)\wedge A'=0$; so by
Lemma~\ref{lem_desired_blade_factorization}, there exists a scalar $\alpha'\in\R$ such that $A'=\alpha'\bigwedge_{i=1}^r p(x_i)$.
It is clear now that
\begin{equation}
A = \frac{\alpha}{\alpha'}A'.
\end{equation}
\end{proof}

The importance of Theorem~\ref{thm_geo_set_rep} can be realized
in the utility of the conformal model.  (See \cite{} for information about the conformal model.)
By Theorem~\ref{thm_geo_set_rep}, we can be justified in algebraically
relating two independently made formulations
of a given geometric set.  For example, we may equate the intersection of two spheres as some
scalar multiple of the canonical intersection of a sphere centered on a plane.  The former formulation
is what we may wish to calculate, while the latter formulation lends itself to analysis through decomposition.
The applicability of Theorem~\ref{thm_geo_set_rep} comes in realizing that each formulation is irreducible.
Reducible blades directly representative of the same geometric set may sometimes be equated as scalar multiples
of one another, but this is not always possible.

The counter-part of Theorem~\ref{thm_geo_set_rep} in algebraic geometry is Hilbert's Nullstellensatz, (see \cite{}),
which implies that for every algebraic set, there exists a unique radical ideal representative of that set.

\subsubsection{Intersections/Unions}

Speaking of algebraic sets, knowing that the intersection or union of any two such sets is also algebraic,
one must wonder if there is an analogous result with regards to geometric sets.  It is easy to find a counter-example
in the conformal model showing that the union of two geometric sets need not be geometric.
Proving or disproving that the intersection of any two geometric sets is geometric, however, is not obvious.
In any case, the following lemma gives us some insight into the union and intersection of geometric sets.

\begin{lem}\label{lem_union_intersect}
For blades $B,C\in\B$, if a non-zero blade $A\in\B$ has a factorization $A=B\wedge C$, then
\begin{align}
\gh(A) &\supseteq \gh(B)\cup\gh(C),\label{equ_geo_set_union} \\
\gd(A) &= \gd(B)\cap\gd(C).\label{equ_geo_set_intersection}
\end{align}
\end{lem}
\begin{proof}
The first of these two equations (equation \eqref{equ_geo_set_union}) being an obvious statement,
we need only show here the validity of equation \eqref{equ_geo_set_intersection}.

If $\bigwedge_{i=1}^s b_i$ is a factorization of the $s$-blade $B$, and
$\bigwedge_{i=1}^t c_i$ is a factorization of the $t$-blade $C$, we have,
by the identity of equation \eqref{equ_vec_blade_inner_prod_series} and Lemma~\ref{lem_lin_indep_blades},
\begin{equation}
\gd(A) = \bigcap_{i=1}^s\gd(b_i)\cap\bigcap_{i=1}^t\gd(c_i) = \gd(B)\cap\gd(C).
\end{equation}
\end{proof}

In light of Lemma~\ref{lem_union_intersect}, we can begin to understand the difficulties found in determining
whether a given intersection of geometric sets is geometric.  Letting $S,T\subset\R^n$ be
a pair of geometric sets, let $B,C\in\B$ be a pair of blades dually representative of them, respectively.
Then, if $A=B\wedge C\neq 0$, we have found, by Lemma~\ref{lem_union_intersect}, a blade $A$ dually
representative of the intersection $\gd(B)\cap\gd(C)$, showing that it is a geometric set.
If, on the other hand, $B\wedge C=0$, we can come to no such conclusion.

\subsubsection{Transformations}

In this section we will take an interest in the set of all versors of $\G$ that preserve the form
of our function $p:\R^n\to\V$.
\begin{defn}
A versor $V\in\G$ with the property that for all $x\in\R^n$, there exists a unique $y\in\R^n$ and
$\gamma\in\R$ such that
\begin{equation}\label{equ_vpv_p}
V^{-1}p(x)V = \gamma p(y),
\end{equation}
is what we will refer to as a {\it preservative versor}.
\end{defn}

The significance of such versors is that the question of how they transform a geometric set
from $\gh(A)$ to $\gh(VAV^{-1})$ is reduced to the question
of how the versor maps points from $\R^n$ to $\R^n$ through our function $p:\R^n\to\V$
as mapping a point $x\in\R^n$ to the point $y\in\R^n$ satisfying equation \eqref{equ_vpv_p}.

The following lemma shows us that in considering the transformation of a geometric set
by a preservative versor, we need only look at the irreducible blades directly representative of that geometric set.

\begin{lem}
Let $V\in\G$ be a preservative versor.
Then for every blade $A\in\B$, if $C\in\B$ is an irreducible blade such that $\gh(A)=\gh(C)$, then
\begin{equation}\label{equ_transformed_reducible_is_transformed_irreducible}
\gh(VAV^{-1})=\gh(VCV^{-1}).
\end{equation}
\end{lem}
\begin{proof}
Letting $A=B\wedge C$ by Lemma~\ref{lem_desired_partial_blade_factorization}, consider the equation
\begin{equation}
VAV^{-1} = VBV^{-1}\wedge VCV^{-1}.
\end{equation}
Notice that $VCV^{-1}$ is an irreducible blade by the preservative property of $V$.
What we must show now is
that $VCV^{-1}$ is, in terms of dimension, the largest irreducible subspace of $VAV^{-1}$.  Supposing for the moment that it isn't,
this would allow us, by the preservative property of $V$, to easily find an irreducible subspace of $A$ that is larger than that of $C$,
which contradicts the fact that $C$ is the largest irreducible subspace of $A$.
\end{proof}

\begin{lem}\label{lem_empty_blade_transforms_as_empty_blade}
Let $V\in\G$ be a preservative versor and $A\in\B$.  Then if $\gh(A)=\emptyset$, we have
\begin{equation}
\gh(VAV^{-1})=\emptyset.
\end{equation}
\end{lem}
\begin{proof}
Showing the contrapositive of our lemma, let $x\in\gh(VAV^{-1})$.  Then, by Lemma~\ref{lem_desired_partial_blade_factorization},
there exists a blade $B\in\B$ such that $VAV^{-1}=B\wedge p(x)$.  We then see that for
some $y\in\R^n$ and a scalar $\gamma\in\R$, we have
\begin{equation}
A = V^{-1}(B\wedge p(x))V = V^{-1}BV\wedge V^{-1}p(x)V = V^{-1}BV\wedge \gamma p(y).
\end{equation}
We now see that $y\in\gh(A)$, showing that $\gh(A)$ is non-empty.
\end{proof}

\subsection{Reduction}

In the special case of a geometric algebra $\G$ having a euclidean signature, (no null vectors),
an algorithm for reducing a given blade $A\in\B$ is given as follows.

\begin{algorithmic}
\Function{Reduce}{$A$}
	\State $x\gets$\Call{FindAnyPoint}{$A$}
	\If{$p(x)\wedge A=0$}
		\State\Return $p(x)\,\wedge\,$\Call{Reduce}{$p(x)^{-1}\cdot A$}
	\EndIf
	\State\Return $1$
\EndFunction
\end{algorithmic}
Here, the function ``FindAnyPoint'' tries to find, (perhaps using methods of calculus), and return any point in $\gh(A)$.
If it fails, the returned point is not in $\gh(A)$.

To see that the function ``Reduce'' above is correct, we first realize that if $p(x)\wedge A=0$,
then there exists a factorization of $A$ of the form $A=p(x)\wedge B$ for some blade $B\in\B$
of grade $r-1$.
Then, by the Gram-Schmidt orthogonalization process, (which is always
applicable in geometric algebras with euclidean signatures), we can rewrite $A$ as a versor,
giving us $A=p(x)B$.  It then follows that
\begin{equation}
p(x)^{-1}\cdot A = \langle p(x)^{-1}p(x) B\rangle_{r-1}=B.
\end{equation}

Notice that the algorithm above is simply a factorization algorithm that is looking for
vector factors of a specific form; and that it may terminate before the blade is fully factored;
in which case, the given blade $A$ was reducible.  A version of this algorithm that works
in any geometric algebra may be adaptable from one given in \cite{}.

\subsection{Example}

Here it may be worth giving an example that illustrates an application of Theorem~\ref{thm_geo_set_rep}.
We begin by letting $\G$ be a geometric algebra generated by a 10-dimensional euclidean vector space $\V$, letting $n=3$, and then defining
$p:\R^n\to\V$ as
\begin{align}
p(x) &= e_0 \nonumber\\
 &+ x_1e_1 + x_2e_2 + x_3e_3 \nonumber\\
 &+ x_2x_3e_4 + x_1x_3e_5 + x_1x_2e_6 \nonumber\\
 &+ x_1^2e_7 + x_2^2e_8 + e_3^2e_9,
\end{align}
where here we're using the notation\footnote{This notation is overloaded.  In some cases, $x_i$ denotes a member
of a sequence of points, while in others, $x_i$ denotes a component of $x$ as a vector.  The meaning of $x_i$ should
always be clear from context.} $x_i=x\cdot e_i$.  (Notice that here we're letting $\R^n$ be
a 3-dimensional subspace of $\V$, and that for all $x\in\R^3$, $p(x)\neq 0$.)  Having done this,
we turn our attention to an equation for the $e_3$-axis-aligned elliptical cylinder, given by
\begin{equation}
\frac{(x_1-h)^2}{a^2}+\frac{(x_2-k)^2}{b^2}-1=0.
\end{equation}
Factoring $p(x)$ out of this equation in terms of the inner product, we get $p(x)\cdot E=0$,
where $E$ is given by
\begin{equation}
E = \left(\frac{h^2}{a^2}+\frac{k^2}{b^2}-1\right)e_0-2\frac{h}{a^2}e_1-2\frac{k}{b^2}e_2+\frac{1}{a^2}e_7+\frac{1}{b^2}e_8.
\end{equation}
It follows that for a non-zero scalar $\alpha\in\R$, $\alpha E\wedge e_3$ is a canonical form for a dually represented ellipse in the $x_3=0$ plane.
Given a blade $A\in\B$ known to dually represent such a shape in this plane, one would hope to decompose it using the following equations.
\begin{align}
h &= (-e_{31}\cdot A)(2e_{37}\cdot A)^{-1}\label{equ_decompose_first} \\
k &= (-e_{32}\cdot A)(2e_{38}\cdot A)^{-1} \\
\lambda &= (h^2e_{37}+k^2e_{38}-e_{30})\cdot A \\
a &= \sqrt{\lambda(e_{37}\cdot A)^{-1}} \\
b &= \sqrt{\lambda(e_{38}\cdot A)^{-1}}\label{equ_decompose_last}
\end{align}
(Here, $e_{ijk\dots}=e_ie_je_k\dots$.)
The problem with this idea is that it can be easily shown that $E\wedge e_3$ is a reducible blade; and therefore,
the blades $A$ and $E\wedge e_3$ may not be scalar multiples of one another.
To see why, consider the following equation which expresses the form of $p(x)$ for all points in the $x_3=0$ plane.
\begin{equation}\label{equ_p_form_in_x3_eq_z_plane}
p(x_1e_1+x_2e_2)=e_0 + x_1e_1 + x_2e_2 + x_1x_2e_6 + x_1^2e_7 + x_2^2e_8
\end{equation}
Now notice that an upper-bound on the dimension of a vector space that can be spanned by vectors
of this form is clearly 6 as there are only 6 components on the right-hand side of equation \eqref{equ_p_form_in_x3_eq_z_plane}.
But $E\wedge e_3$ is a 2-blade, making its dual a blade of grade $10-2=8>6$.  It follows that there is no set of
8 points $\{x_i\}_{i=1}^8$ on an ellipse for which $\bigwedge_{i=1}^8 p(x_i)\neq 0$.

To illustrate the case that $A$ is not a scalar multiple of $E\wedge e_3$, consider the solution set
to the equation
\begin{equation}
x_1^2 + x_2^2 - (x_3+1)^2\tan\frac{\pi}{4}=0,
\end{equation}
which gives us a canonical surface partially submerged below the $x_3=0$ plane.
Factoring $p(x)$ out of this equation in terms of the inner product, we get $p(x)\cdot C=0$,
where $C$ is given by
\begin{equation}
C = -e_0 + 2e_3 + e_7 + e_8 - e_9.
\end{equation}
Letting $A=C\wedge e_3$ and seeing that $A$ dually represents a circle in the $x_3=0$ plane, let us consider
the problem of decomposing it into its characteristic parts using equations
\eqref{equ_decompose_first} through \eqref{equ_decompose_last}.
Again, it is easy to see that $AI$ is reducible, and further, that it is no scalar multiple of $E\wedge e_3$.
(Note that $I=e_{0123456789}$.)
Fully reducing $AI$, we get the 5-blade $C'$, given by
\begin{equation}
C' = e_{01267} - e_{01268} + e_{12678} = e_{126}\wedge(-e_{07} + e_{08} + e_{78}).
\end{equation}
Then, by Theorem~\ref{thm_geo_set_rep}, this must be a subspace of $(E\wedge e_3)I$.
Sure enough, we have
\begin{equation}
(e_{459}\wedge C')I = E\wedge e_3,
\end{equation}
and we get the unit circle in the $x_3=0$ plane, centered at origin.

The lesson learned here is that we can make use of canonical forms if we first
fully reduce the blade we wish to decompose.  Sometimes, however, this is undesirable.
For example, if we were to fully reduce a blade directly representative of a tangent
point in the conformal model, then we would be left with an ordinary point and lose the tangential information.
This brings up the point that a blade may embed within it more information than just the geometric
set that it represents dually or directly.

\section{Closing}

In closing, a few remarks might be made.  First, there are surely more results to be had
about geometric sets in an abstract setting.  The advantage to seeking out such
results is that they will be applicable to any model of geometry we invent
in geometric algebra that uses blades to represent subsets of $\R^n$ as we have
done in this paper.  On the other hand,
we may find that more advanced models of geometry use multi-vectors generally
for such a purpose, or something other than geometric algebra altogether.

Lastly, there was some hope that stronger connections between geometric algebra
and algebraic geometry would be found in this paper.  There has even been some
lofty and naive expectation among geometric algebraists, (the present author not included),
that a reformulation of algebraic geometry in terms of geometric algebra would somehow enrich the subject
in ways beyond the current capabilities of abstract algebra.  However, even with a limited understanding
of algebraic geometry, it is abundantly clear that this notion is absurd.  Nevertheless,
geometric algebra should not be discounted; for it may bring some contributions to the study of solution sets
of algebraic equations.

\begin{thebibliography}{9}

\end{thebibliography}

\end{document}