\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

\title{Chapter 15 Exercises\\Gallian's Book on Abstract Algebra}
\author{Spencer T. Parkin}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{identity}{Identity}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{result}{Result}[section]

%\newcommand{\gcd}{\mbox{gcd}}
\newcommand{\lcm}{\mbox{lcm}}
\newcommand{\abs}{\mbox{abs}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\G}{\mathbb{G}}
\newcommand{\stab}{\mbox{stab}}
\newcommand{\aut}{\mbox{Aut}}
\newcommand{\inn}{\mbox{Inn}}
\newcommand{\orb}{\mbox{orb}}
\newcommand{\chr}{\mbox{char}\,}

\begin{document}
\maketitle

\section*{Exercise 1}

Prove Theorem 15.1.

For the theorem, we let $\phi$ be a ring homomorphism from a ring $R$ to a ring $S$,
and we let $A$ be a subring of $R$ and let $B$ be an ideal of $S$.

The first part of the theorem states that for any $r\in R$ and any positive integer $n$,
that $\phi(nr)=n\phi(r)$ and $\phi(r^n)=(\phi(r))^n$.

Proof: That $\phi(nr)=n\phi(r)$ follows from Property 2 of Theorem 10.1 for group homomorphisms.
Similarly,
\begin{equation*}
\phi(r^n)=\phi(\underbrace{r\cdot\dots\cdot r}_n)=\underbrace{\phi(r)\cdot\dots\cdot\phi(r)}_n=(\phi(r))^n.
\end{equation*}

The second part of the theorem states that $\phi(A)=\{\phi(a)|a\in A\}$ is a subgring of $S$.

Proof: That $\phi(A)$ is an Abelian group follows from Properties 1 and 3 of Theorem 10.2
for group homomorphisms.  Then, if $a,b\in\phi(A)$, then there exist $x,y\in A$ such
that $\phi(x)=a$ and $\phi(y)=b$.  Then, since $xy\in A$ and $\phi(xy)=\phi(x)\phi(y)=ab$,
we see that $ab\in\phi(A)$.  Having now shown closure of the ring multiplication of $S$
in $\phi(A)$, we can claim that $\phi(A)$ is a subring of $S$.

The third part of the theorem states that if $A$ is an ideal and $\phi$ is onto $S$,
then $\phi(A)$ is an ideal.

Proof: By the second part of this theorem, $\phi(A)$ is a subring, so we need
only prove now that it is an ideal of $S$.  Let $s\in S-\phi(A)$ and $y\in\phi(A)$.  Then since $\phi$ is onto, there
exists $r\in R$ such that $\phi(r)=s$.  Let $x\in A$ such that $\phi(x)=y$.
Then since $rx\in A$, (because $A$ is an ideal of $R$), and $\phi(rx)=\phi(r)\phi(x)=sy$,
we have $sy\in\phi(A)$.  Similarly, since $xr\in A$, (again, because $A$ is an ideal of $R$),
and $\phi(xr)=\phi(x)\phi(r)=ys$, we have $ys\in\phi(A)$.  We can now claim that $\phi(A)$
is an ideal of $S$.

The fourth part of the theorem states that $\phi^{-1}(B)=\{r\in R|\phi(r)\in B\}$ is an ideal of $R$.

Proof: By Property 7 of Theorem 10.2, $\phi^{-1}(B)$ is a subgroup of $R$.  It must be an
Abelian group since all subgroups of rings are Abelian.  Now let $r\in R$ and $x\in\phi^{-1}(B)$.
Then $\phi(r)\in S$ and $\phi(x)\in B$ and since $B$ is an ideal of $S$, $\phi(rx)=\phi(r)\phi(x)\in B$,
showing that $rx\in\phi^{-1}(B)$.  Similar reasoning shows that $xr\in\phi^{-1}(B)$, so $\phi^{-1}(B)$
is an ideal of $R$.

The fifth part of the theorem states that if $R$ is commutative, then $\phi(R)$ is commutative.

Proof: Letting $a,b\in R$, notice that
\begin{equation*}
\phi(a)\phi(b)=\phi(ab)=\phi(ba)=\phi(b)\phi(a).
\end{equation*}

The sixth part of the theorem states that if $R$ has a unity 1, $S\neq\{0\}$, and $\phi$ is onto,
then $\phi(1)$ is the unity of $S$.

Proof: Notice that for all $r\in R$, we have $\phi(1)\phi(r)=\phi(r)$.  Then since $\phi$ is onto,
it follows that $\phi(1)s=s$ for all $s\in S$.  This shows that $\phi(1)$ is either the unity of $S$,
or that $\phi(r)=0$ for all $r\in R$.  Now if $S\neq\{0\}$ and $\phi$ is onto, then we can't
have $\phi(r)=0$ for all $r\in R$.  So $\phi(1)=1$.

The seventh part of the theorem states that $\phi$ is an isomorphism if and only if $\phi$ is onto and
$\ker\phi=\{r\in R|\phi(r)=0\}=\{0\}$.

Proof: This follows immediately from Property 9 of Theorem 10.2.  We need only look at the statement
from a purely group-theoretic stand-point and also realize that $\phi$ will preserve the multiplication
product of the ring.

The eighth and last part of the theorem states that if $\phi$ is an isomorphism from $R$ onto $S$,
then $\phi^{-1}$ is an isomorphism from $S$ onto $R$.

Proof: Realize that $\ker\phi^{-1}$ is the trivial subring of $R$.  This part of the theorem then
follows from the seventh part of the theorem.

\section*{Exercise 2}

Prove Theorem 15.2.

Let $\phi$ be a homomorphism from a ring $R$ to a ring $S$.  Then
$\ker\phi=\{r\in R|\phi(r)=0\}$ is an ideal of $R$.

Proof: From group theory, we already know that $\ker\phi$ is
a normal subgroup of $R$.  Now let $r\in R$ and $x\in\ker\phi$.
Then $\phi(rx)=\phi(r)\phi(x)=\phi(r)\cdot 0=0\implies rx\in\ker\phi$.
Similarly, we have $xr\in\ker\phi$, so $\ker\phi$ is an ideal of $R$.

\section*{Exercise 3}

Prove Theorem 15.3.

Let $\phi$ be a ring homomorphism from $R$ to $S$.  Then the
mapping from $R/\ker\phi$ to $\phi(R)$, given by $r+\ker\phi\to\phi(r)$,
is an isomorphism.  In symbols, $R/\ker\phi\approx\phi(R)$.

Proof: By Theorem 10.3, $\phi$ is a group isomorphism from $R/\ker\phi$
to $\phi(R)$.  Now since $\ker\phi$ is an ideal, $R/\ker\phi$ is a factor ring
by Theorem 14.2.  What remains to be shown is that the mapping preserves
multiplication in $\phi(R)$.  To that end, see that for any pair of elements $x,y\in R$, we have
\begin{equation*}
\Psi(x+\ker\phi)\Psi(y+\ker\phi)=\phi(x)\phi(y)=\phi(xy)=\Psi(rs+\ker\phi),
\end{equation*}
where $\Psi:R/\ker\phi\to\phi(R)$ is the mapping given in the theorem's statement.

\section*{Exercise 4}

Prove Theorem 15.4.

Every ideal of a ring $R$ is the kernel of a ring homomorphism of $R$.
In particular, an ideal $A$ is the kernel of the mapping $r\to r+A$ from
$R$ to $R/A$.

Proof: Define $\phi(r)=r+A$ as the natural homomorphism from $R$ to $R/A$.
It is not hard to see that $\phi$ preserves both operations of $R$ in $R/A$.
Clearly, $\phi(r)=A$ if and only if $r\in A$, so $\ker\phi=A$.

\section*{Exercise 18}

Determine all ring isomorphisms from $Z_n$ to itself.

We know that all such isomorphisms $\phi:Z_n\to Z_n$ are of
the form $\phi(x)=x\phi(1)$.  Then, by Property 6 of Theorem
15.1, we must have $\phi(1)=1$.  So $\phi(x)=x$ is the
only isomorphism from $Z_n$ to itself.

\section*{Exercise 24}

Recall that a ring element $a$ is called an idempotent if $a^2=a$.
Prove that a ring homomorphism carries an idempotent to an idempotent.

Let $\phi$ be a ring homomorphism and let $a$ be an idempotnent
in the domain of $\phi$.  Then $\phi(a)^2=\phi(a^2)=\phi(a)$
by Property 1 of Theorem 15.1.

\section*{Exercise 36}

Determine all ring homomorphisms from $Q$ to $Q$.

By Exercise 40 of Chapter 6, all such homomorphisms are
of the form $\phi(x)=x\phi(1)$.  (A ring homomorphism must also
be a group homomorphism.)  Furthermore, since we must
have $\phi(1)=\phi(1\cdot 1)=\phi(1)\phi(1)$, $\phi(1)$ must be
an idempotent of $Q$.  But the only idempotents are 0 and 1.
So $\phi(x)=x$ and $\phi(x)=0$ are the only two homomorphisms
of $Q$ to itself.

\section*{Exercise 48}

Suppose that $n$ divides $m$ and that $a$ is an idempotent of $Z_n$ (that is,
$a^2=a$).  Show that the mapping $x\to ax$ is a ring homomorphism from $Z_m$
to $Z_n$.  Show that the same correspondence need not yield a ring homomorphism
if $n$ does not divide $m$.

Let $\phi:Z_m\to Z_n$ be defined as $\phi(x)=ax$ for an idempotent $a$ of $Z_n$.
In considering whether $\phi$ is well defined, we have to ask ourselves: for $x,y\in Z_m$,
if $x\equiv y\pmod m$, then do we have $\phi(x)\equiv\phi(y)\pmod n$?  Well, if $n|m$ and
$m|(x-y)$, then $n|(x-y)\implies n|a(x-y)=ax-ay$.  So $\phi$ is well defined.
We can now go on to show the $\phi$ preserves addition and multiplication.
We have
\begin{equation*}
\phi(x+y)=a(x+y)=ax+ay=\phi(x)+\phi(y),
\end{equation*}
and since $a$ is an idempotent of $Z_n$, we have
\begin{equation*}
\phi(xy)=axy=a^2xy=axay=\phi(x)\phi(y).
\end{equation*}
Suppose now that $m=2$ and $n=3$.  Let $a=1$.
Notice that while $0\equiv 2\pmod 2$, we have
$\phi(0)=0\not\equiv2=\phi(2)\pmod 3$.

\section*{Exercise 53}

Let $D$ be an integral domain and let $F$ be the field of quotients of $D$.
Show that if $E$ is any field that contains $D$, then $E$ contains a subfield
that is ring-isomorphic to $F$.  (Thus, the field of quotients of an integral
domain $D$ is the smallest field containing $D$.)

Let $\phi:F\to E$ be a function defined as $\phi(a/b)=ab^{-1}$, where $a,b\in D$ with $b\neq 0$.
To see that this is a well defined function, let $a',b'\in D$ such that $a/b=a'/b'$.
It follows that $ab'=a'b$, so $ab^{-1}=a'b(b')^{-1}b^{-1}=a'(b')^{-1}$, showing
that $\phi$ is indeed well defined.  Now let $x,y\in D$ with $y\neq 0$ and
see that
\begin{align*}
\phi\left(\frac{a}{b}+\frac{x}{y}\right)&=\phi\left(\frac{ay+bx}{by}\right) \\
 &=(ay+bx)(by)^{-1} \\
 &= ab^{-1} + xy^{-1} \\
 &= \phi\left(\frac{a}{b}\right)+\phi\left(\frac{x}{y}\right),
\end{align*}
showing that $\phi$ preserves addition.  We then see that
\begin{equation*}
\phi\left(\frac{ax}{by}\right)=(ax)(by)^{-1}=ab^{-1}xy^{-1}=\phi\left(\frac{a}{b}\right)\phi\left(\frac{x}{y}\right),
\end{equation*}
showing that $\phi$ preserves multiplication.  It follows that $\phi$ is
a homomorphism from $F$ to $E$.

Now see that $\ker\phi=\{0/1\}$, since if $ab^{-1}=0$ and $b^{-1}\neq 0$, we must have $a=0$.
Then since $\phi$ is clearly onto $\phi(F)$, it follows from Property 7 of
Theorem 15.1 that $\phi(F)$ is ring-isomorphic to $F$.  Now realize that $\phi(F)$ is a
subfield of $E$.

\end{document}